---
title: "20230530 Client Summary"
author: "Stephanie Hickey"
date: "5/30/2023"
output: html_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(hrbrthemes)
library(lubridate)
clients = read.delim("20230530_client_summary.txt")
```

Since August 2022, we have had `r nrow(clients)` client requests, including requests for analysis advice, grant support, and data analysis services, from `r length(unique(clients$PI))` different research groups across `r length(unique(clients$Institution))` institutions.

### Requests by request service type
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=3.5, fig.height=3.5, fig.align='center'}
tally_service = 
  clients %>% 
  group_by(Service.Type) %>%
  tally()

p = 
  ggplot(tally_service, aes(x=Service.Type, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), vjust=-0.3, size=3.5) +
    ylab("number of requests") +
    theme_classic()
  
plot(p)

ggsave(p, file = "requests_per_service_type.png", width = 3.5, height = 3.5)

n_charged = nrow(clients %>%
                         filter(!is.na(Charge)))
```

We have invoiced or sent estimates for `r n_charged` data analysis projects, totaling `r paste0("$",sum(clients$Charge, na.rm = TRUE))` with a median cost of `r paste0("$",median(clients$Charge,, na.rm = TRUE))` per project.

### Requests per project type
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width = 4, fig.height = 5,fig.align='center'}
tally_service = 
  clients %>% 
  select(Project.Type.Broad) %>%
  mutate(Project.Type.Broad = strsplit(as.character(Project.Type.Broad), ",")) %>% 
  unnest(Project.Type.Broad) %>%
  group_by(Project.Type.Broad) %>%
  tally()

tally_service$Project.Type.Broad <- factor(tally_service$Project.Type.Broad,                  
                          levels = tally_service$Project.Type.Broad[order(tally_service$n, decreasing = TRUE)])

p = 
  ggplot(tally_service, aes(x=Project.Type.Broad, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), vjust=-0.3, size=3.5) +
    ylab("number of requests") +
    xlab("Project Type") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
  
plot(p)

ggsave(p, file = "requests_per_analysis_type.png", width = 7, height = 5)
```

### MSU requests per department and college

Requests come from `r length(clients %>% filter(Institution == "MSU") %>% pull(Deparment) %>% unique)` departments across `r length(clients %>% filter(Institution == "MSU") %>% pull(College) %>% unique)` colleges.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=15}
tally_dep = 
  clients %>% 
  filter(Institution == "MSU") %>%
  group_by(Deparment) %>%
  tally() %>%
  arrange(desc(n))

tally_dep$Deparment <- factor(tally_dep$Deparment,                  
                          levels = tally_dep$Deparment[order(tally_dep$n, decreasing = FALSE)])

p1 =
ggplot(tally_dep, aes(x=Deparment, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), hjust=-0.3, size=3.5) +
    ylab("number of requests") +
    xlab("Department") +
    theme_classic() +
    coord_flip()
    #theme(axis.text.x = element_text(angle = 45, hjust=1))

tally_col = 
  clients %>% 
  filter(Institution == "MSU") %>%
  group_by(College) %>%
  tally() %>%
  arrange(desc(n))

tally_col$College <- factor(tally_col$College,                  
                          levels = tally_col$College[order(tally_col$n, decreasing = FALSE)])

p2 =
ggplot(tally_col, aes(x=College, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), hjust=-0.3, size=3.5) +
    ylab("number of requests") +
    xlab("College") +
    theme_classic() +
    coord_flip()

p = p1 + p2
plot(p)
ggsave(p, file = "requests_per_department_college.png", width = 15, height = 8)
```

### Requests over time
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=5}

clients$Contact_Date = mdy(clients$Contact_Date)

tally_date = 
  clients %>%
  mutate(month = lubridate::floor_date(Contact_Date, "month")) %>%
  group_by(month) %>%
  tally()

p <- ggplot(tally_date, aes(x=month, y=n)) +
  geom_line(color="#18453B") +
  geom_point() +
  ylab("number of requests") +
  scale_x_date(date_labels="%b %y", date_breaks  ="1 month") +
theme_minimal()
  
p

ggsave(p, file = "requests_per_month.png", width = 9, height = 5)
```