---
title: "FY23.Q1 - FY25.Q4 Summary"
author: "Stephanie Hickey"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    smaller: yes
    widescreen: yes
---

<style type="text/css">

h2 {
  text-align: left;
  position: fixed;
  top: 50%;
  left: 50%;
  transform: translate(-50%, -50%);
}
}
</style>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork)
library(hrbrthemes)
library(lubridate)
library(ggalluvial)

clients = read.csv("/Users/stephanielepp/Desktop/BioinformaticsCore/business_summaries/clients.csv")

invoiced = 
  readxl::read_excel("/Users/stephanielepp/Desktop/BioinformaticsCore/business_summaries/invoiced.xlsx") %>%
  rename(Total.Cost = "Total Cost",
         Invoice.Date = "Invoice Date")

```

# 

### Requests by service type

Since August 2022, we have had `r nrow(clients)` client requests, including requests for analysis advice, grant support, and data analysis services, from `r length(unique(clients$PI))` different research groups across `r length(unique(clients$Institution))` institutions.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=3.5, fig.align='center'}
tally_service = 
  clients %>% 
  group_by(Service.Type) %>%
  tally()

p = 
  ggplot(tally_service, aes(x=Service.Type, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), vjust=-0.3, size=3.5) +
    ylab("number of requests") +
    theme_classic()
  
plot(p)

ggsave(p, file = "requests_per_service_type.png", width = 5, height = 3.5)
```

We have invoiced `r nrow(invoiced)` data analysis projects, totaling `r paste0("$", sum(invoiced$Total.Cost, na.rm = TRUE))` with a median cost of `r paste0("$", median(invoiced$Total.Cost, na.rm = TRUE))` per project.

# 
### Requests per project type for project types with $>1$ request.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#fig.width = 4, fig.height = 5,fig.align='center'
tally_service = 
  clients %>% 
  select(Project.Type.Broad) %>%
  mutate(Project.Type.Broad = gsub(", ", ",", Project.Type.Broad)) %>% 
  mutate(Project.Type.Broad = strsplit(as.character(Project.Type.Broad), ",")) %>% 
  unnest(Project.Type.Broad) %>%
  group_by(Project.Type.Broad) %>%
  tally()

tally_service$Project.Type.Broad <- factor(tally_service$Project.Type.Broad,                  
                          levels = tally_service$Project.Type.Broad[order(tally_service$n, decreasing = TRUE)])

p = 
  tally_service %>%
  filter(n > 1) %>%
  ggplot(aes(x=Project.Type.Broad, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), vjust=-0.3, size=3.5) +
    ylab("number of requests") +
    xlab("Project Type") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
  
plot(p)

ggsave(p, file = "requests_per_analysis_type.png", width = 7, height = 5)
```

# 
### MSU requests per department and college

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# {r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=15}

clients =
  clients %>%
  rename(Department = "Deparment") %>%
  mutate(Department = trimws(Department),
         PI = trimws(PI)) %>%
  mutate(Department = gsub("Department of ", "", Department)) %>%
  mutate(College = gsub("College of ", "", College)) %>%
  mutate(Department = 
           ifelse(str_detect(Department, "Biochemistry"), 
                  "Biochemistry & Molecular Biology", Department)) %>%
  mutate(Department = 
           ifelse(str_detect(Department, "Molecular Genetics"), 
                  "Microbiology, Genetics, & Immunology", Department)) %>%
  mutate(Department = 
           ifelse(str_detect(Department, "Pathobiology"), 
                  "Pathobiology and Diagnostic Investigation", Department)) %>%
  mutate(Department = 
           ifelse(str_detect(PI, "Kevin Childs"), 
                  "Genomics Core", Department),
         Department = 
           ifelse(str_detect(PI, "Elena Demireva"), 
                  "Transgenic Core", Department),
         College = 
           ifelse(str_detect(Department, "Core"), 
                  "RTSF", College),
         PI = 
           ifelse(str_detect(PI, "Erin Percell"), 
                  "Erin Purcell", PI),
         College = 
           ifelse(str_detect(Department, "Plant, Soil and Microbial Sciences"), 
                  "CANR", College),
         College = 
           ifelse(str_detect(Department, "Microbiology, Genetics, & Immunology"), 
                  "CNS", College),
         College = 
           ifelse(str_detect(Department, "Physiology"), 
                  "Human Medicine", College))

n_dep = length(clients %>% 
                 filter(Institution == "MSU",
                        !is.na(Department)) %>%
                 pull(Department) %>% 
                 unique)          

n_college = length(clients %>% 
                     filter(Institution == "MSU", !is.na(College)) %>% 
                     pull(College) %>% unique)
```

Requests come from `r n_dep` departments across `r n_college` colleges.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4.5, fig.width=7.5}
tally_dep = 
  clients %>% 
  filter(Institution == "MSU",
         !is.na(Department)) %>%
  group_by(Department) %>%
  tally() %>%
  arrange(desc(n))

tally_dep$Department <- factor(tally_dep$Department,                  
                          levels = tally_dep$Department[order(tally_dep$n, decreasing = FALSE)])

p1 =
ggplot(tally_dep, aes(x=Department, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), hjust=-0.3, size=2) +
    ylab("number of requests") +
    xlab("Department") +
    theme_classic() +
    coord_flip()
    #theme(axis.text.x = element_text(angle = 45, hjust=1))

tally_col = 
  clients %>% 
  filter(Institution == "MSU",
          !is.na(Department)) %>%
  group_by(College) %>%
  tally() %>%
  arrange(desc(n))

tally_col$College <- factor(tally_col$College,                  
                          levels = tally_col$College[order(tally_col$n, decreasing = FALSE)])

p2 =
ggplot(tally_col, aes(x=College, y=n)) +
    geom_bar(stat="identity", fill = "#18453B") +
    geom_text(aes(label=n), hjust=-0.3, size=3.5) +
    ylab("number of requests") +
    xlab("College") +
    theme_classic() +
    coord_flip()

p = p1 + p2

tally_river = 
  clients %>% 
  filter(Institution == "MSU",
         !is.na(Department)) %>%
  group_by(Department, College) %>%   # Group by College and Department
  tally(name = "department_count") %>%  # Count occurrences of each department, name the count column
  ungroup() %>%
  group_by(College) %>%
  mutate(college_count = sum(department_count)) %>%  # Add total count per college
  ungroup() %>%
  mutate(College = paste0(College," (", college_count, ")"),
         college_count = 
           ifelse(str_detect(College, "Human Medicine"), 
                 college_count + .5 , college_count))

tally_river = tally_river[with(tally_river, order(college_count, department_count, College)),]

tally_river$Department = factor(tally_river$Department, levels = tally_river$Department)

tally_river$College = factor(tally_river$College, levels = rev(unique(tally_river$College)))

msu_pallet = c("#18453B", "#49CFA5", "#066B00", "#7bbd00", "#089000", "#044c36", "#8CFF98", "#C3FFEC", "#87FFD9", "#0B9A6D", "#077351")

ggplot(tally_river, aes(x=Department, 
                        y=department_count, 
                        fill = College)) +
    geom_bar(stat="identity") +
    scale_fill_manual(values = msu_pallet) +
    geom_text(aes(label=department_count), hjust=-0.3, size=2) +
    ylab("number of requests") +
    xlab("Department") +
    theme_classic() +
    theme(text = element_text(size = 10)) +
    coord_flip() 
ggsave(file = "requests_per_department_college.png", width = 15, height = 10)
```

# 
### Requests over time

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#, fig.width=9, fig.height=5

clients$Contact_Date = mdy(clients$Contact_Date)

tally_date = 
  clients %>%
  #mutate(month = lubridate::floor_date(Contact_Date, "month")) %>%
  #group_by(month) %>%
  mutate(Quarter = lubridate::quarter(Contact_Date, type = "year.quarter", fiscal_start = 7)) %>%
  group_by(Quarter) %>%
  tally()

# p <- ggplot(tally_date, aes(x=Quarter, y=n)) +
#   geom_line(color="#18453B") +
#   geom_point() +
#   ylab("number of requests") +
#   scale_x_date(date_labels="%b %y", date_breaks  ="1 month") +
# theme_minimal()

p <- ggplot(tally_date, aes(x=as.factor(Quarter), y=n, group=1)) +
  geom_line(color="#18453B") +
  geom_smooth(method = "lm", se=FALSE, color="#7BBD00") +
  geom_point(size=1) +
  geom_label(label = tally_date$n, size = 2.5) + 
  ylab("number of requests") +
  xlab("fiscal quarter") +
theme_minimal()
  
p

ggsave(p, file = "requests_per_month.png", width = 9, height = 5)
```

# 
### Requests over time by service type

```{r, echo=FALSE, message=FALSE, warning=FALSE}

tally_date = 
  clients %>%
  #mutate(month = lubridate::floor_date(Contact_Date, "month")) %>%
  #group_by(month) %>%
  mutate(Quarter = lubridate::quarter(Contact_Date, type = "year.quarter", fiscal_start = 7)) %>%
  group_by(Quarter, Service.Type) %>%
  tally()

ggplot(tally_date, aes(x = as.factor(Quarter), y = n, fill = Service.Type, label = n)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = msu_pallet[1:5]) +
  geom_text(size = 3, position = position_stack(vjust = 0.5)) +
  theme_minimal() +
  xlab("Quarter") +
  ylab("Requests")

```

# 
### Total amount invoiced over time by quarter

```{r, echo=FALSE, eval=T, message=FALSE}
tally_inv = 
  invoiced %>%
  #mutate(month = lubridate::floor_date(Contact_Date, "month")) %>%
  #group_by(month) %>%
  mutate(Quarter = lubridate::quarter(Invoice.Date, type = "year.quarter", fiscal_start = 7)) %>%
  group_by(Quarter) %>%
  summarise(Total.Invoiced = sum(Total.Cost)) %>%
  mutate(Curency = paste('$',formatC(Total.Invoiced, big.mark=',', format = 'f', digits = 2)))

p <- ggplot(tally_inv, aes(x=as.factor(Quarter), y=Total.Invoiced, group=1)) +
  geom_line(color="#18453B") +
  geom_smooth(method = "lm", se=FALSE, color="#7BBD00") +
  geom_point(size=1) +
  geom_label(label = tally_inv$Curency, size = 2.5) + 
  ylab("Total Invoiced") +
  xlab("fiscal quarter") +
theme_minimal()
  
p
```

# 

### Total amount invoiced over time by fiscal year

```{r, echo=FALSE, eval=T, message=FALSE}
tally_inv = 
  invoiced %>%
  #mutate(month = lubridate::floor_date(Contact_Date, "month")) %>%
  #group_by(month) %>%
  mutate(Quarter = lubridate::quarter(Invoice.Date, type = "year.quarter", fiscal_start = 7),
         FY = stringr::str_sub(Quarter, 1, 4)) %>%
  group_by(FY) %>%
  summarise(Total.Invoiced = sum(Total.Cost)) %>%
  mutate(Curency = paste('$',formatC(Total.Invoiced, big.mark=',', format = 'f', digits = 2)))

p <- ggplot(tally_inv, aes(x=as.factor(FY), y=Total.Invoiced, group=1)) +
  geom_line(color="#18453B") +
  geom_smooth(method = "lm", se=FALSE, color="#7BBD00") +
  geom_point(size=1) +
  geom_label(label = tally_inv$Curency, size = 2.5) + 
  ylab("Total Invoiced") +
  xlab("fiscal year") +
theme_minimal()
  
p
```

#

### Total projects invoiced over time by fiscal year

```{r, echo=FALSE, eval=T, message=FALSE}
tally_inv = 
  invoiced %>%
  mutate(Quarter = lubridate::quarter(Invoice.Date, type = "year.quarter", fiscal_start = 7),
         FY = stringr::str_sub(Quarter, 1, 4)) %>%
  group_by(FY) %>%
  tally(name = "Total.Invoiced")
 
p <- ggplot(tally_inv, aes(x=as.factor(FY), y=Total.Invoiced, group=1)) +
  geom_line(color="#18453B") +
  geom_smooth(method = "lm", se=FALSE, color="#7BBD00") +
  geom_point(size=1) +
  geom_label(label = tally_inv$Total.Invoiced, size = 2.5) + 
  ylab("Total Projects Invoiced") +
  xlab("fiscal year") +
theme_minimal()
  
p
```

#

### Office hours total visits per month  

```{r, echo=FALSE, eval=T, message=FALSE}
office_hours =
  read.csv("/Users/stephanielepp/Desktop/BioinformaticsCore/business_summaries/Bioinformatics_Core_Help_Hours.csv") %>%
  filter(!is.na(Name))

office_hours$Date = mdy(office_hours$Date)

tally_oh = 
  office_hours %>%
  mutate(month = lubridate::floor_date(Date, unit = "month"),
         month = format(as.Date(month), "%Y-%m")) %>%
  group_by(month) %>%
  tally() 

p <- ggplot(tally_oh, aes(x=as.factor(month), y=n, group=1)) +
  geom_line(color="#18453B") +
  geom_smooth(method = "lm", se=FALSE, color="#7BBD00") +
  geom_point(size=1) +
  geom_label(label = tally_oh$n, size = 2.5) + 
  ylab("Number of Visits") +
  xlab("Month") +
theme_minimal()
  
p
```

Over `r length(unique(tally_oh$month))` months, we have had `r length(unique(office_hours$Name))` unique visitors to our office hours.